Adaptive group-regularized logistic elastic net regression

### 2019-02-08 docs/manuscript.Rnw
added second simulation

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,mathtools,bbm,pgfplotstable,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,pgfplotstable,threeparttable, booktabs,mathtools,dsfont,parskip}

@@

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, VU University Medical Center, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands\\
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\
	3. MRC Biostatistics Unit, Cambridge, United Kingdom \\

@@

	A natural way to deal with (b) is through differential penalization: that is, each group of variables receives its own penalty parameter. An apparent issue with this differential penalization is the estimation of the penalty parameters. Naive estimation may be done by cross-validation. However, cross-validation requires re-estimation of the model over a grid of penalty parameters. The size of this grid increases exponentially with the number of penalty parameters. Consequently, it quickly becomes computationally infeasible. We therefore propose an efficient alternative: empirical Bayes estimation of the penalty parameters, which corresponds to hyperparameter estimation in the Bayesian prior framework. Because of the ubiquity of binary outcome data in omics research, we focus on the logistic elastic net. Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in spirit. Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection.

@@

	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as a computationally more convenient scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$. Using this result, we write the generalised elastic net model in its Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i \bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this Bayesian formulation the penalty parameters in $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	We simulated data according to five different scenarios: 
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations within groups of features, 
		\item differential signal between the groups and heavy-tailed distributed model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 10 of the SM. In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform. In the second scenario, we expect \texttt{gren} and \textt{GRridge} to overfit due to the large number of hyperparameters to estimate. The rest of the methods are conjectured to perform better, due to a lack of overfitting. In scenario three, we expect all methods to perform badly due to the high correlation between the features. In scenario three

@@

	 The results of this simulation are presented in Figures \ref{fig:sim1_performance1_set1}-\ref{fig:sim1_performance1_set3}.
	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set1.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 1.}
		\label{fig:sim1_performance1_set1}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set2.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 1.}
		\label{fig:sim1_performance1_set2}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim1_res1_performance1_set3.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 1.}
		\label{fig:sim1_performance1_set3}
	\end{figure}

	The results of this simulation 1 are presented in Figures \ref{fig:sim4_performance1_set1}-\ref{fig:sim4_performance1_set3}.
	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set1.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.05$ based on simulation scenario 4.}
		\label{fig:sim4_performance1_set1}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set2.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.5$ based on simulation scenario 4.}
		\label{fig:sim4_performance1_set2}
	\end{figure}

	\begin{figure}[!h]
		\centering
		\includegraphics[width=1\linewidth]{gren_sim4_res1_performance1_set3.png}
		\caption{LOESS fit of (a) Cohen's kappa, (b) MSE, (c) AUC, and (d) BSS for ridge, \texttt{GRridge}, elastic net, \texttt{gren}, \texttt{SGL}, \texttt{grMCP}, and \texttt{gel} with $\alpha=0.95$ based on simulation scenario 4.}
		\label{fig:sim4_performance1_set3}
	\end{figure}

>>>>

\documentclass[a4paper,hidelinks]{article}
\usepackage{url,bm,algorithm,algpseudocode,bbm,bm,amsmath,amssymb,amsthm,amsfonts,graphics,graphicx,epsfig,rotating,caption,subcaption,natbib,appendix,titlesec,multicol,hyperref,verbatim,bbm,algorithm,algpseudocode,threeparttable, booktabs,dsfont,parskip}

@@

	\noindent
	1. Department of Epidemiology \& Biostatistics, Amsterdam Public Health research institute, Amsterdam University medical centers, PO Box 7057, 1007 MB
	Amsterdam, The Netherlands \\*
	2. Mathematical Institute, Leiden University, Leiden, The Netherlands \\*
	3. MRC Biostatistics Unit, Cambridge University, Cambridge, United Kingdom

@@

	A natural way to deal with (b) is through differential penalization: that is, each group of variables receives its own penalty parameter. An apparent issue with this differential penalization is the estimation of the penalty parameters. Naive estimation may be done by cross-validation. However, cross-validation requires re-estimation of the model over a grid of penalty parameters. The size of this grid increases exponentially with the number of penalty parameters. Consequently, it quickly becomes computationally infeasible. We therefore propose an efficient alternative: empirical Bayes estimation of the penalty parameters, which corresponds to hyperparameter estimation in the Bayesian prior framework. Because of the ubiquity of binary outcome data in omics research, we focus on the logistic elastic net. Recently, \cite{zhang_novel_2018} introduced VBVS, a method similar to ours in spirit. Nevertheless, one clear advantage of our method over VBVS is the possibility of adaptive inclusion of external information on the features to aid in prediction and feature selection. We would like to emphasize that our method is different from recent developments in grouped penalized regression such as \cite{dondelinger_joint_nodate}, where the grouping is over the observations as opposed to the features

@@

	\cite{li_bayesian_2010} show that (\ref{eq:elasticnetprior}) may be written as a computationally more convenient scale mixture of normals, with mixing parameter $\btau = \begin{bmatrix} \tau_1 & \cdots & \tau_p \end{bmatrix} \tr$. Using this result, we write the generalised elastic net model in its Bayesian form as:
	\begin{subequations}\label{eq:enmodel2}
		\begin{align}
		\y| \bbeta & \sim \prod_{i=1}^n \mathcal{B} \left( m_i, \expit ( \x\tr_i \bbeta ) \right),  \\
		\bbeta | \btau & \sim \prod_{g=1}^G \prod_{j \in \G(g)} \mathcal{N} \left(0,\frac{1}{\lambda' _g \lambda_2} \frac{\tau_j - 1}{\tau_j} \right), \\
		\btau & \sim \prod_{j=1}^p \mathcal{TG} \left( \frac{1}{2},\frac{8 \lambda_2}{\lambda_1^2}, \left(1,\infty \right) \right).
		\end{align}
	\end{subequations}
	Here, $\mathcal{TG} \( k,\theta,\( x_l,x_u\) \)$ denotes the truncated gamma distribution with shape $k$, scale $\theta$, and domain $\(x_l,x_u\)$. In this Bayesian formulation the penalty parameters $\blambda = \begin{bmatrix} \lambda_1 & \lambda_2 & \lambda'_1 & \cdots & \lambda'_G \end{bmatrix} \tr$ play the role of the hyperparameters in a Bayesian hierarchical model.

@@

	We simulated data according to five different scenarios:
	\begin{enumerate}
		\item differential signal between the groups and uniformly distributed model parameters,
		\item a large number of small groups of features,
		\item no differential signal between the groups, but strong correlations within groups of features,
		\item differential signal between the groups and heavy-tailed distributed model parameters, and
		\item a very sparse setting with no signal in some of the groups.
	\end{enumerate}
	Full descriptions of the scenarios are given in Section 10 of the SM. In the first scenario, the distribution of the model parameters is far off from the assumed ridge/elastic net/lasso distributions, so we assume that all methods will underperform. In the second scenario, we expect \texttt{gren} and \texttt{GRridge} to overfit due to the large number of hyperparameters to estimate. The rest of the methods are conjectured to perform better, due to a lack of overfitting. In scenario three, we expect all methods to perform badly due to the high correlation between the features. In scenario three

@@
